---
title: "Analysis of aggregate data"
author: "Zerihun Bekele"
date: "8/21/2019"
output: 
  html_document:
    code_folding: show
    keep_md: true
    toc: true
    toc_depth: 3
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import libraries

```{r message=FALSE}
library("mi")
library("dplyr")
```

## Read static data

```{r}
static_data <- read.csv("Analysis/static_agg_data.csv", header=T, row.names = 1)
print(dim(static_data))
static_data[1:10, 1:4]
```

### Number of cases with PI 

The NA values in the `pi` and `pi onset date` variables imply pressure injury was not observed or a negative result, so we have to convert the NAs to 0 and the rest to 1.

```{r fig.width=6, fig.asp=0.7}

print(c("Number of cases with PI = ", sum(!is.na(static_data$pi))), quote = FALSE)

# Replace NA with 0 and convert to factor
#static_data$pi <- pi.data$pi
static_data$pi[is.na(static_data$pi)] <- 0
static_data$pi_dayfromadmit[is.na(static_data$pi_dayfromadmit)] <- Inf

static_data$pi <- factor(static_data$pi, levels=c(0,1), labels = c("No", "Yes"))

# Calculate percent pi and non pi
pi.rate <- round(prop.table(table(static_data$pi))*100,2)

# Plot chart
{par(mai=c(0.5,0.6,0.5,0.2))
barplot(pi.rate, xlim = c(0,2),width = 0.75, ylim = c(0,100), col = c("tomato", "green3"), main = "Cases with PI, %")
text(y = pi.rate, x = c(0.6,1.5), labels = paste(pi.rate, '%'), pos=3, xpd = T, adj = 0.5, cex = 1.2, offset = 0.3)
box()}
```

### Missing values in the static data 

```{r}
na.count <- function(df) apply(is.na(df), 2, sum )
```

```{r fig.height=12, fig.width=8}
na.static <- na.count(static_data)

# Sort and compute percent missing
na.static <- sort(na.static[na.static > 0]/nrow(static_data)*100)

clrs = circlize::colorRamp2(c(0,90),c("green","red"), 0.2)(1:100)

# Define a function to plot barplot
na.barplot <- function(df) {
  par(mai=c(0.5, 2.8, 0.4,0.2))
  xbar <- barplot(df, xlim = c(0, max(df)*1.2),
                ylim = c(1,length(df)*1.2),
                names.arg = F, border = NA, horiz = T, 
                main = "Percent missing", 
                col = clrs[as.integer(df+1)])
  text(x= -5, y=xbar, labels = names(df), xpd=T, 
     pos= 2, adj = 1, cex=0.9, offset = -0.5)
  text(x= df, y=xbar, labels = round(df,1), 
     xpd=T, pos= 4, adj = 1, cex=0.8, offset = 0.5)
  box()}

# Plot barplot for static missing values
na.barplot(na.static)
```

### Get rows with at least one missing feature

```{r fig.width=10, fig.asp=0.8}
na.row.idx <- apply(is.na(static_data), 1, any)
na.rows <- static_data[na.row.idx, ]
print(dim(na.rows))

z <- apply(na.rows, 2, function(x) ifelse(is.na(x), 0, 1))

par(mai=c(1.5,0.8,0.5,0.2), oma = c(3,0.5,0.1, 0.1))
image(1:ncol(z), 1:nrow(z),t(z), xaxt='n',
        ylab = "Individuals", xlab = "",
        main = "Individuals with at least one feature missing")
axis(1, 1:ncol(z), labels = NA, tck = -0.02)
text(1:ncol(na.rows), -nrow(na.rows)/25, labels = names(na.rows), srt = 90, 
       adj = 0, xpd = T, pos = 2, cex = 0.8, offset = 0)
mtext("Note: The pale yelow color indicates missing values.",outer = T, side = 1, line = 1,
        cex = 0.8,adj = 0, col = "red")
```

## Read time series binary data

```{r}
ts_bi_data <- read.csv("Analysis/ts_binary_agg.csv", header=T, row.names = 1)
print(dim(ts_bi_data))
ts_bi_data[1:15,1:5]
```

### Checking missing values

```{r}
na.ts_bi <- na.count(ts_bi_data)

# Compute percent missing
na.ts_bi <- sort(na.ts_bi[na.ts_bi > 0]/nrow(ts_bi_data)*100)

# Plot barplot for static missing values
print(na.ts_bi)
```


## Read time series numeric data

```{r}
ts_num_data <- read.csv("Analysis/ts_numeric_agg.csv", header=T, row.names = 1)
print(dim(ts_num_data))
ts_num_data[1:15, 1:5]
```

### Checking missing values

```{r fig.asp=2}
na.ts_num <- na.count(ts_num_data)

# Compute percent missing
na.ts_num <- sort(na.ts_num[na.ts_num > 0]/nrow(ts_num_data)*100)

# Plot barplot for static missing values
na.barplot(na.ts_num)
```

```{r}
get.na.rows <- function(df) {
  na.row.idx <- apply(is.na(df), 1, any)
  na.rows <- df[na.row.idx, ]
  print(dim(na.rows))
  par(mai=c(1.5,0.8,0.5,0.2), oma = c(3,0.5,0.1, 0.1))
  image(1:ncol(na.rows), 1:nrow(na.rows), t(as.matrix(na.rows)), xaxt='n',
        ylab = "Individuals", xlab = "",
        main = "Individuals with at least one feature missing")
  axis(1, 1:ncol(df), labels = NA, tck = -0.02)
  text(1:ncol(na.rows), -nrow(na.rows)/20, labels = names(na.rows), srt = 90, 
       adj = 0, xpd = T, pos = 2, cex = 0.8, offset = 0)
  mtext("Note: white color indicates missing values.",outer = T, side = 1, line = 1,
        cex = 0.8,adj = 0, col = "red")
}
```


## Read braden aggregate data

```{r}
braden_data <- read.csv("Analysis/braden_agg_perc.csv", header=T, row.names = 1)
print(dim(braden_data))
braden_data[1:15, 1:5]
```

### Checking missing values

```{r fig.asp=2}
na.braden <- na.count(braden_data)

# Compute percent missing
na.braden <- sort(na.braden[na.braden > 0]/nrow(braden_data)*100)

# Plot barplot for static missing values
na.barplot(na.braden)
```

```{r fig.width=8, fig.asp=0.7}
get.na.rows(braden_data)
```

The above plot indicates that all the missing rows are correlated.

## Read periop numeric data

```{r}
periop_num_data <- read.csv("Analysis/periop_numeric_agg.csv", header=T, row.names = 1)
print(dim(periop_num_data))
periop_num_data[1:15, 1:4]
```

### Checking missing values

```{r fig.width=6, fig.asp=1.5}
na.periop_num <- na.count(periop_num_data)

# Compute percent missing
na.periop_num <- sort(na.periop_num[na.periop_num > 0]/nrow(periop_num_data)*100)

# Plot barplot for static missing values
na.barplot(na.periop_num)
```

```{r fig.width=8, fig.asp=0.7 }
get.na.rows(periop_num_data)
```

## Read periop binary (yes/no) data

```{r}
periop_bi_data <- read.csv("Analysis/periop_yn_agg.csv", header=T, row.names = 1)
print(dim(periop_bi_data))
periop_bi_data[1:10, 1:4]
```

### Checking missing values

**Here it is not clear whether there is a missing value or there was no use of the specific type of treatment.**

```{r fig.width=6, fig.asp=1.5}
na.periop_bi <- na.count(periop_bi_data)
print(any(na.periop_bi>0))
```

## Read location data

```{r}
location_data <- read.csv("Analysis/location_agg.csv", header=T, row.names = 1)
print(dim(location_data))
location_data[1:10,1:4]
```

### Checking missing values

```{r fig.width=6, fig.asp=1.5}
na.location <- na.count(location_data)
print(any(na.location>0))

# Compute percent missing
na.location <- sort(na.location[na.location > 0]/nrow(location_data)*100)

# Plot barplot for static missing values
na.barplot(na.location)
```

## Read surgical data

```{r}
surgical_data <- read.csv("Analysis/surgical_data_agg.csv", header=T, row.names = 1)
print(dim(surgical_data))

surgical_data[1:10,1:4]

unique(gsub("_\\d+$","", names(surgical_data)))
```

**We need to figure out what each the values in these variables mean**

### Checking missing values

```{r fig.width=6, fig.asp=2}
na.surgical <- na.count(surgical_data)
print(any(na.surgical>0))

# Compute percent missing
na.surgical <- sort(na.surgical[na.surgical > 0]/nrow(surgical_data)*100)

# Plot barplot for static missing values
na.barplot(na.surgical)
```

## Select features with NA less than 30%

```{r}
all.data <- cbind(static_data, ts_bi_data,ts_num_data,braden_data,
                  periop_bi_data,periop_num_data, location_data, surgical_data)

all_vars_na <- na.count(all.data)
na.perc <- sort(all_vars_na/nrow(all.data)*100)

vars_30 <- names(na.perc)[which(na.perc <= 30,)]

all.data.30 <- all.data[,vars_30]
write.csv(all.data.30, "all.data.30.csv")

xx <- hist(na.perc, breaks = 10, col = "palegreen", ylim = c(0,200))
text(xx$breaks+5, xx$counts, labels = xx$counts, pos = 3)

```

```{r}
library(DT)

datatable(all.data.30[sample(nrow(all.data.30),100),])
```


## Select features with NA less than 10%

```{r}
vars_10 <- names(na.perc)[which(na.perc <= 10,)]

all.data.10 <- all.data[,vars_10]

```

## Analyze & Impute missing data for static data type

```{r message=FALSE, fig.width=12, fig.height=8}
library("betareg"); library("mi")

# Get variables with 30 pct less missing values and drop pi and pi day from onset features
static.na <- na.count(static_data[, -which(names(static_data) %in% c("pi", "pi_dayfromadmit"))])*100/nrow(static_data)
static_data.30pct <- static_data[, names(static.na)[which(static.na <= 30,)]]

# get show the missing information matrix
mdf <- missing_data.frame(static_data.30pct) 

image(mdf)
```

```{r}
show(mdf)
```

### Start imputation

```{r message=FALSE}
#imputations <- mi(mdf, n.iter = 5, n.chains = 3, max.minutes = 10, verbose = TRUE)
#show(imputations)
#summary(static_data.30pct)
library(mice)
system.time(imputed_Data <- mice(static_data.30pct, m=5, maxit = 30, seed = 500, printFlag = F))
save(imputed_Data, file = "static_imputed_Data.RData")
#md.pattern(static_data.30pct)
summary(imputed_Data)

densityplot(imputed_Data)
```
```{r}
library(dplyr)
library(caret)
static_imp1 <- complete(imputed_Data, 1)
#names(static_imp1)[sapply(static_imp1, is.factor)]

cat_data <- select_if(static_imp1, is.factor)
```

Imputation using the `missForest' package.

```{r}
library(missForest)
library(doParallel)
registerDoParallel(cores=3)
system.time(static_imputed <- missForest(xmis = static_data.30pct, maxiter = 10, parallelize = 'variables', ntree = 50))
```







