---
title: "PIMPS PressureInjury"
author: "SOCR COllaboration (Dinov-Anderson-Tschannan)"
date: "May 10, 2019"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("G:/My Drive/PIPM_Data_R_code")
```

## Overview

[GDrive](https://drive.google.com/drive/folders/15A2ytB_foR6pwlQ2M6imbjsPDnrxNdYl)

This project involves a large, high-dimensional and time series dataset related to pressure injuries (bed ulcers). Specifically, your thoughts on segmentation/clustering/classification of this data are of interest. Also cross-validation strategy thoughts would help us. 

## Aims

#### Aim 1: Develop and validate a predictive model of PI risk among hospitalized surgical patients. 

 - Identify clusters associated with PI risks for hospitalized surgical patients.
 
 - Develop and validate a predictive model of PI risk tailored to the highest PI risk clusters. 	

#### Aim 2: Conduct a pilot study to determine the implementation and subsequent impact of a dynamic, personalized PI prevention intervention informed by the predictive model from Aim 1 on hospital acquired PIs among surgical patients.

## Import data

```{r import data, echo=T, message=F}
library("memisc")
system.time({
  data <- as.data.set(spss.system.file("July_2019_Data_Docs/PIPM_6-24-19.sav"))
  pi.data <- as.data.frame(data)
  })

dim(pi.data)

#write.table(pi.data, "PIPM_cases.txt", sep = "\t")
#PI.data <- read.table("PIPM_cases.txt", header=T, sep = "\t")
# [1]   100 11180

# 30
# 40
# < 40
# Supervised and unsupervised model
# Clusters
# Hierarichical clustering
# 9 o'clock next firday 23rd

```

## Get the static data

Next, separate the static data, which is located from column 1 to 58.

```{r message=F}
library(dplyr)
static_data <- pi.data[,1:58]

dim(static_data)
```

Add aditional static data and remove some redundent variables from the dataframe.

```{r remove.date.vars}
#Remove non relevant variables
rm_vars <- c("encounterid", "admitdt", "deceased_date", "pi_dateonset","termcodemapped_group", "icu_stay", "termc0", "dischargedt")

static_data <- subset(static_data, select = !(colnames(static_data) %in% rm_vars))
static_data <- cbind(static_data, dplyr::select(pi.data, dplyr::starts_with("lab")))

write.csv(static_data, "Analysis/static_agg_data.csv")

dim(static_data)
```
Print the variable names in the static data.

```{r print.col.names, eval=FALSE}
# Print variable names
for (i in seq(length(static_data)) ) {
  print(c(colnames(static_data)[i], class(static_data[,i])))
  
}
```

### Number of cases with PI

```{r}

print(c("Number of cases with PI = ", sum(!is.na(static_data$pi))), quote = FALSE)

# Replace NA with 0 and convert to factor
static_data$pi <- pi.data$pi
static_data$pi[is.na(static_data$pi)] <- 0

static_data$pi <- factor(static_data$pi, levels=c(0,1), labels = c("No", "Yes"))

# Calculate percent pi and non pi
pi.rate <- round(prop.table(table(static_data$pi))*100,2)

# Plot chart
{par(mai=c(0.5,0.6,0.5,0.2))
barplot(pi.rate, xlim = c(0,2),width = 0.75, ylim = c(0,100), col = c("tomato", "green3"), main = "Cases with PI, %")
text(y = pi.rate, x = c(0.6,1.5), labels = paste(pi.rate, '%'), pos=3, xpd = T, adj = 0.5, cex = 1.2, offset = 0.3)
box()}
```


## Generate summary statistics 

```{r print.summary}
for (i in 1:length(static_data)) {
  print(colnames(static_data)[i])
  print(summary(static_data[,i]))
  
}

```

## Generate simple plots

```{r race_ethnicity, fig.width=10, fig.height=8}
library(ggplot2)

ggplot(static_data, aes(x = race, fill = ethnicity)) +
   geom_bar() +
   theme(axis.text.x = element_text(angle = 15, hjust = 1, size=10))

```

```{r age_los, echo=FALSE, fig.height=6, fig.width=10}

ggplot(static_data, aes(x = ageinyears, y = los)) +
   geom_point(aes(color = admissiontype), size=2) +
   theme(axis.text.x = element_text(size=12))

ggplot(static_data, aes(x = los, color=factor(pi, labels = c("No", "Yes")))) +
  geom_density(size=1) +
  labs(title = "Length of stay", color = "PI")+
  theme(legend.position = c(0.9,0.9))+
  scale_color_manual(values=c("tomato","yellowgreen"))

```

```{r bar_hist_plot}
#Let us generate a histogram of a barplot to see what the data looks like

bar_hist <- function (dt) {
  n.row <- ceiling(length(dt)/5)
  par(mfrow=c(n.row,5), cex=0.7, cex.main=1.2,
      mai=c(0.3,1,0.3,0.25), 
      mar=c(2.5,1.8,2.5,1.5))
  
  for (i in seq(length(dt)) ) {
  
    #Get the column
    column.i <- dt[,i]
    ttle <- colnames(dt)[i]
    
    # We want to plot barplot if the column/feature is a categorical type
    bar = is.factor(column.i)
    #Plot bar plot if it is a factor variable
    if (bar) {
      x.prop <- prop.table(table(column.i))
      barplot(x.prop, xpd=FALSE, main = ttle, 
              col = "#CC79A7", 
              ylim = c(0,max(x.prop)+0.1))
      
    }
    
    #Plot a density if it is a numeric
    else {
      #plot(density(column.i, na.rm = T), lwd=2, main = ttle)
      #hist(column.i, main =colnames(x)[i] )
      boxplot(column.i, horizontal = T)
    }
  }

}

```

```{r plots, fig.width=14, fig.height=10}
# Use the above fucntion to plot the data
bar_hist(static_data[,1:25])
bar_hist(static_data[,26:50])
bar_hist(static_data[,51:69])

```

## Bivariate plots

```{r mosaic.plot}
#Plot for all data set
library(vcd)
(PGA <- structable(admissiontype ~ pressure_injury_poa + gender, data = static_data))
mosaic(PGA, shade = TRUE, legend=TRUE)

# Plot only for PI =Yes

pi.data <- subset(static_data, !is.na(pi))
(pi.PGA <- structable(admissiontype ~ pressure_injury_poa + gender, data = pi.data))
mosaic(pi.PGA, shade = TRUE, legend=TRUE)

pairs(PGA, space = 0.15)
```

```{r}

hist(static_data$pi_dayfromadmit, 
     col = "tan2", 
     main = "PI development day from admission",
     xlab = "PI date onset",
     freq = F)

hist(static_data$los, 
     col = "tan2", 
     main = "Length of stay",
     xlab = "Day",
     freq = F)
```


```{r fig.width=8}
ggplot(static_data, aes(x = los, y = pi_dayfromadmit)) +
  geom_point(aes(color=obesity, alpha=0.2, size=weightkg))+
  guides(alpha="none")

```

**Note: Negative days for PI day onset doesn't make sense.**

```{r fig.width=10}
clr1 = c( "tomato2", "orange", "steelblue2")

ggplot(static_data) + 
  geom_histogram(aes(x = ageinyears, fill = admissiontype), color = 'grey', bins = 7) +
  facet_wrap(~ gender) +
  ggtitle("For all data")+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_fill_brewer(palette="Set1")

ggplot(pi.data) + 
  geom_histogram(aes(x = ageinyears, fill = admissiontype), color = 'grey', bins = 7) +
  facet_wrap(~ gender) +
  ggtitle("For PI positive data")+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_fill_brewer(palette="Set1")
```

```{r}
#par(mfrow=c(1,2))
ggplot(static_data) + 
  geom_point(aes(x = ageinyears, y = los, color=pi, size=as.integer(los)/2, alpha=0.3)) +
  guides(alpha="none")+
  scale_size_continuous(name = "Length \nof stay")

ggplot(static_data) + 
  geom_point(aes(x = los, y = bmi, size = ageinyears, alpha=0.2)) +
  guides(alpha="none")
```
```{r}
ggplot(static_data, aes(x = pi, y= weightkg))+
  geom_boxplot()

#ggplot(static_data, aes(x = weightkg, color = pi))+
#  geom_density()
```
## Check missing values in the static data

```{r fig.height=12, fig.width=8}
na.count <- apply(apply(static_data, 2, is.na), 2, sum)
na.count <- sort(na.count[na.count > 0]/nrow(static_data)*100)


{
par(mai=c(0.5, 2.8, 0.4,0.2))
clrs = circlize::colorRamp2(c(0,90),c("green","red"), 0.2)
xbar <- barplot(na.count, xlim = c(0, max(na.count)+10), 
                names.arg = F, border = NA, horiz = T, 
                main = "Percent missing", 
                col = clrs(0:length(na.count)))
text(x= -5, y=xbar, labels = names(na.count), xpd=T, 
     pos= 2, adj = 1, cex=0.9, offset = -0.5)
text(x= na.count, y=xbar, labels = round(na.count,1), 
     xpd=T, pos= 4, adj = 1, cex=0.8, offset = 0.5)
box()}
```
## Get timeseries data

```{r ts_data, message=FALSE}
library(tidyverse)
ts_data <- dplyr::select(pi.data, starts_with('day'))
var.names <- colnames(ts_data)
#var.names <- gsub("\\d+", "", var.names)
var.names <- gsub("day\\d+\\_","", var.names)
unique.vars <- unique(var.names)
dim(ts_data)
unique.vars
```
### Compute summaries for the daily data

#### Aggregate binary features

Select binary features and compute summaries.

First, let us compute the percent of time a treatment was given over the length of stay (LOS).

```{r}
#Binary features
bi_ts_vars <- c("steroid", "vasopressor", "parenteral_nutrition", "intubated",
                "incontinent", "fecaldevice", "potubefeed")

# Empty data frame to hold the results
ts_bi_perc <- as.data.frame(matrix(data=NA, nrow = nrow(ts_data), ncol = length(bi_ts_vars)))
colnames(ts_bi_perc) <- paste(bi_ts_vars, "perc", sep = "_")

# Table for yes or no
ts_bi_use <- ts_bi_perc
colnames(ts_bi_use) <- paste(bi_ts_vars, "used", sep = "_")

# Compute percent: # days with treatment/# los in hospital
treatment_percent <- function (x) (apply(x, 1, sum, na.rm=T)*100/pi.data$los)

for (i in 1:length(bi_ts_vars)) {
  # Select a specific treatment type one by one
  trt.dt <- dplyr::select(ts_data, ends_with(bi_ts_vars[i]))
  
  # Calculate percent of los with each type of treatment
  ts_bi_perc[,i] <- treatment_percent(trt.dt)
  
  # Check if there is use of a specific treatment type.
  ts_bi_use[,i] <- apply(!is.na(trt.dt), 1, any)
}

summary(ts_bi_perc)

summary(ts_bi_use)

# Combine the time series binary data
ts_bi_agg <- cbind(ts_bi_use, ts_bi_perc)

# Write to file
write.csv(ts_bi_agg, "Analysis/ts_binary_agg.csv")

```
#### Plot distribution

```{r binary.perc.plot, fig.height=6 }
par(mfrow=c(3,2), oma = c(0.5,1,5,1), mai=c(0.5,0.6,0,0))

for (i in 1:6) {
  xx <- hist(ts_perc[,i], main ="", freq = F, col = "red", breaks = 10, border = NA, xlab = "")
  text(50, max(xx$density), bi_ts_vars[i], cex = 1.5, pos = 1 )
  box()
}
mtext(" Histograms of percent of LOS with a specific treatment", side = 3, line = 2, outer = T, cex = 1.3, adj = 0)

```


#### Compute summaries for numeric features

Most of these features have a daily min and max value. What we want is the average of both. Therefore, we are going to calculate the mean of the max and min values of all features for each individual. This will generate a dataframe with rows equal to the number of subjects and colums equal to the number of unique features, then we can plot the distribution of the means.

```{r ts.numeric.aggregation}
# Get non-binary data variable names
non_bi_vars <-setdiff(unique.vars, bi_ts_vars)

# get the numeric variables
ts_num_vars <- grep("braden", non_bi_vars, value = T, invert = T)

# get variables for minimum values
ts_num_min_vars <- grep("min", ts_num_vars, value = T)

# table to hold absolute min value for each variable
ts_min_tbl <- data.frame(matrix(NA, nrow = nrow(ts_data), ncol = length(ts_num_min_vars)))
colnames(ts_min_tbl) <- ts_num_min_vars

# Get variables with max values
ts_num_max_vars <- grep("max", ts_num_vars, value = T)

# table to hold absolute min value for each variable
ts_max_tbl <- data.frame(matrix(NA, nrow = nrow(ts_data), ncol = length(ts_num_max_vars)))
colnames(ts_max_tbl) <- ts_num_max_vars

# Compute absolute minimum for the numeric variables
get.min <- function(x) ifelse(all(is.na(x)), NA, min(x, na.rm = T))

for (i in ts_num_min_vars) {
  min_tbl <- select(ts_data, contains(i))
  ts_min_tbl[[i]] <- apply(min_tbl, 1, get.min)
}

# Compute absolute maximum for the numeric variables
get.max <- function(x) ifelse(all(is.na(x)), NA, max(x, na.rm = T))

for (i in ts_num_max_vars) {
  max_tbl <- select(ts_data, contains(i))
  ts_max_tbl[[i]] <- apply(max_tbl, 1, get.max)
}


##############################################################
# Empty tables to hold the results for mean values
ts_num_mean <- data.frame(matrix(data=NA, nrow = nrow(ts_data), ncol = length(ts_num_vars)))
colnames(ts_num_mean) <- ts_num_vars

# Compute the mean of min and max for all hospital stay period
avg_value <- function (x) (apply(x, 1, mean, na.rm=T))

for (i in ts_num_vars) {
  # Select one type of treatment/feature at a time
  select_dt <- select(ts_data, contains(i))
  
  # Compute the mean across length of stay for each individual
  ts_num_mean[[i]] <- avg_value(select_dt)
}
colnames(ts_num_mean) <- paste0(ts_num_vars, "_avg")

###################################################
# Combine all numeric aggregate tables to one table
ts_num_agg <- cbind(ts_min_tbl, ts_max_tbl, ts_num_mean)

write.csv(ts_num_agg, "Analysis/ts_numeric_agg.csv")
```

# Check completeness of the numeric aggregate table

```{r fig.width=7, fig.height=14}
num.agg <- apply(ts_num_agg, 2, function(x) sum(!is.na(x))*100/length(x))

{ par(mai=c(0.5, 1.7,0.2,0.2),oma = c(0.5,1,3,1))
  xbar <- barplot(num.agg, horiz = T, xlim = c(0, 100),
                  ylim = c(1,length(num.agg)+6), col = "palegreen3",
                names.arg = F, border = NA)
  text(x= -5, y=xbar, labels = names(num.agg), xpd=T, 
     pos= 2, adj = 1, cex=0.9, offset = 0.1)
  text(x= num.agg, y=xbar, labels = round(num.agg,1),
       xpd=T, pos= 2, adj = 1, cex=0.8, offset = 0.5)
  box()
mtext("Percent completeness of the numeric aggregate dataframe", 
      side = 3, line = 1, outer = T, cex = 1.3, adj = 0)
}

```

#### Aggregate Braden matrics variables

Next, aggregation will be done for each braden matrics variables based patient state.

```{r braden.aggregate}
# get the numeric variables
ts_braden_vars <- grep("braden", non_bi_vars, value = T)

braden_agg <- data.frame(matrix(NA, nrow = nrow(ts_data), ncol = 4*length(ts_braden_vars)))

# Generate column names for each braden activity and levels
colnames(braden_agg) <- unlist(lapply(ts_braden_vars, function (x) paste0(rep(x,4), 1:4)))

# Function to get count for each level
level.counts <- function (x) table(x)[c("1","2","3","4")]

for (i in 1:length(ts_braden_vars)) {
  braden.tbl <- select(ts_data, contains(ts_braden_vars[i]))
  
  k <- 4*(i-1)+1
  
  braden_agg[,k:(k+3)] <- data.frame(t(apply(braden.tbl, 1, level.counts)))
  
}
```

##### Check data completeness

```{r fig.width=7, fig.height=14}
# Check completeness of the columns
braden.cols <- apply(braden_agg, 2, function(x) sum(!is.na(x)))

{ par(mai=c(0.5, 2,0.2,0.2))
  xbar <- barplot(braden.cols, horiz = T, xlim = c(0, max(braden.cols)+200),
                  ylim = c(1,56),
                names.arg = F, border = NA)
  text(x= -5, y=xbar, labels = names(braden.cols), xpd=T, 
     pos= 2, adj = 1, cex=0.9, offset = 0.5)
  text(x= 1500, y=xbar, labels = round(braden.cols,1),
       xpd=T, pos= 4, adj = 1, cex=0.8, offset = 0.5)
  box()}
```
The NA values in the above aggregate table does not indicate missing value unless all are NA per metric type, but it means the individual didn't have that kind of activity level throughout their lOS, thus we can replace it with zero. Then, we will normalize the dataframe by converting it into the percent of times an individual that many number of a specific activity over thier LOS.

```{r }
# We can drop columns with zero counts.
braden_agg <- braden_agg[, braden.cols != 0]

# Calculate the percent of los with a specific activity test
# First set all NAs to zero
braden_agg_perc <- apply(braden_agg, 2, function (x) ifelse(is.na(x), 0, x))

braden_agg_perc <- as.data.frame(apply(braden_agg_perc, 2, function (x) 100*x/pi.data$los), col.names = names(braden_agg))

write.csv(braden_agg_perc, "Analysis/braden_agg_perc.csv")

# Replace back to NA if all row values are zero
# First identify which rows are all zeros
braden.nas <- c()
all.nas.idx <- c()
for (i in ts_braden_vars) {
  select.tbl <- select(braden_agg_perc, contains(i))
  na.rows <- apply(select.tbl, 1, sum) == 0
  braden_agg_perc[na.rows, names(select.tbl)] <- NA
  braden.nas <- append(braden.nas, sum(na.rows))
  all.nas.idx <- union(all.nas.idx, which(na.rows))
}

# The number of missing values in each category
names(braden.nas) <- ts_braden_vars
print(braden.nas)

```

```{r fig.width=10, fig.height=14}
braden_index <- gsub("braden","", names(braden_agg_perc))
par(mfrow=c(12,4), oma = c(0.5,1,4,1), mai=c(0.2,0.2,0.1,0.1))
for (i in 1:ncol(braden_agg_perc)) {
  xx <- hist(braden_agg_perc[,i], main="", freq=F, breaks = 10, border = NA, col = "red", xlab="")
  text(50, max(xx$density), braden_index[i], cex = 1.3, pos = 1 )
  box()
}
mtext(" Histograms of percent of LOS vs percent individual with a specific braden activity level", side = 3, line = 2, outer = T, cex = 1.3, adj = 0)

```

#### Analyze missing data in the Braden dataframe

```{r fig.width=12, fig.height=8}
library("betareg"); library("mi")

# get show the missing information matrix           
mdf <- missing_data.frame(braden_agg_perc) 
#show(mdf)
image(mdf)
```

#### Analyze missing data for timeseries aggregate dataframe

```{r fig.width=12, fig.height=8}
# Combine all time series aggregate tables
ts_agg_combined <- cbind(ts_bi_perc, ts_bi_use, ts_num_agg, braden_agg_perc)
# get show the missing information matrix           
mdf <- missing_data.frame(ts_agg_combined) 
#show(mdf)
image(mdf)

```


```{r}
# Check pi after removing NAs
table(!is.na(static_data$pi[-all.nas.idx]))
```

## Agregation of individual time series data

### Location data

```{r get_location_data}
library(dplyr)

loc_data <- dplyr::select(pi.data, starts_with('location'))

dim(loc_data)

# Calculate length of stay in each locations
?reshape2::melt
locs_los <- dplyr::select(loc_data, contains('enddate')) - dplyr::select(loc_data, contains('startdate'))
#Convert seconds to days
locs_los <- locs_los/86400

colnames(locs_los) <- paste0('loc_', seq(1,15), '_los')

locs_los <- as.data.frame(lapply(locs_los, as.numeric))
head(locs_los)

# Compute the number of individuals in each locations
loc.stay <- apply(!is.na(locs_los), 2, sum)

# Compute number of transfers for each individual
loc.trans <- apply(!is.na(locs_los), 1, sum)
locs_los["transfers"] <- loc.trans

loc_names <- dplyr::select(loc_data, -ends_with('date'))

all_names <- c()
for (i in 1:15) {
  all_names <- c(all_names, unique(as.character(loc_names[,i])))
}

loc_data["rm_locs"] <- (loc_names$location_1 %in% c("MH 10E PED ICU","MH 8E ACUTE CARE","UH 9C ADULT PSYCH"))
write.csv(loc_data, "Analysis/location_data.csv")


sum(loc_data$rm_locs)
write.csv(locs_los, "Analysis/location_agg.csv")
# Top 15 transfers
sort(loc.trans, decreasing = T)[1:15]

# Number of individuals with no location info
sum(loc.trans == 0)

ddd <- apply(loc_data[loc_data$rm_locs,4:45],1, function(x) all(is.na(x)))
sum(ddd)

```

#### Correct los

```{r}
loc_data <- read.csv("Analysis/location_data.csv")
all_pi.data <- read.csv("Analysis/all_piData.csv")
all_pi.data['rm_locs'] <- loc_data$rm_locs
summary(all_pi.data$los)

all_pi.data$los <- as.numeric(as.Date(pi.data$dischargedt) - as.Date(pi.data$admitdt))
write.csv(all_pi.data, "Analysis/all_piData.csv")

remove_locs = c("UH 4A", "UH 4AS STROKE", "UH 6D CRIT CARE ICU", "UH 7C MED ACUTE CARE",
                "UH 7DN CORONARY CARE", "UH 8A3 Onc Acute Care", "UH BAC BURN ACUTE", "UH SHORT STAY UNIT")

locs_to_remove <- loc_data$location_1 %in% remove_locs
sum(locs_to_remove)

table(all_pi.data$pi[!locs_to_remove])
all_pi.data["locs_to_remove"] <- locs_to_remove

write.csv(all_pi.data, "Analysis/all_piData.csv")

```


#### Convert location data to long format

```{r}
id <- as.character(pi.data$encounterid)
loc_data_long <- cbind(id, as.matrix(loc_data[,1:3]))
dim(loc_data_long)

for (i in seq(4,45,3)) {
  mat <- cbind(id,as.matrix(loc_data[,i:(i+2)]))
  
  loc_data_long <- rbind(loc_data_long, mat)
  print(dim(loc_data_long))
  
}
loc_data_long <- data.frame(loc_data_long)
loc_data_long$id <- as.character(loc_data_long$id)
```


```{r}

library(reshape2)
loc_data_long$location_1startdate <-  as.Date(loc_data_long$location_1startdate)

loc_data_long$location_1enddate <-  as.Date(loc_data_long$location_1enddate)
loc_data_long <- loc_data_long[,1:4]

#loc_data_wide <- dcast(loc_data_long, 
#                      id ~ location_1, 
#                       fun.aggregate = function(x) as.Date(x[1]), 
#                       value.var ="location_1startdate")

loc_data_wide2<- reshape(data=loc_data_long, timevar="location_1",
                         idvar = "id",
                         v.name=c("location_1startdate","location_1enddate"),
                         direction="wide")


location_los <- dplyr::select(loc_data_wide2, contains('enddate')) - dplyr::select(loc_data_wide2, contains('startdate'))

location_los <- as.data.frame(apply(location_los,2, function(x) as.numeric(sub(" days","", x))))

location_los[is.na(location_los)] = 0


# save both dataframe to file
names(location_los) <- sub("location_1enddate.", "", names(location_los))
location_los['id'] <- id
write.csv(location_los, "Analysis/location_los.csv", row.names = FALSE)

names(loc_data_wide2) <- sub("location_1", "", names(loc_data_wide2))
write.csv(loc_data_wide2, "Analysis/loc_dates_wide.csv", row.names = FALSE)
```

```{r}
locs.rm <- sub("startdate.", "", names(loc_data_wide2))
locs.rm <- sub("enddate.", "", names(loc_data_wide2))
save_names <- names(loc_data_wide2)
locs.rm <- (locs.rm[2:61] %in% c("MH 10E PED ICU","MH 8E ACUTE CARE","UH 9C ADULT PSYCH"))
names(loc_data_wide2) <- locs.rm

new_loc_data <- select(loc_data_wide2, -matches( "MH 10E PED ICU|MH 8E ACUTE CAR|UH 9C ADULT PSYCH"))
sum(apply(new_loc_data,1, function(x)  sum(!is.na(x))) < 1)
```




## Agreggate periop data


### Get yes/no (binary) periop data

```{r}
periop_vars <- grep("periop", names(pi.data), value = T)
#periop_data <- gsub("periop\\d+\\_","", periop_data)
periop_vars <- gsub("\\d+\\_","_", periop_vars)
periop_vars <- unique(gsub("\\d+$","", periop_vars))
print(periop_vars)
```

```{r}
or_yn_vars <- c("periop_steroid_", "periop_vasopressor_")
or_yn_vars <- append(or_yn_vars, grep("yn$", periop_vars, value = TRUE))

# Empty data frame to hold the results
or_yn_perc <- as.data.frame(matrix(data=NA, nrow = nrow(ts_data), ncol = length(or_yn_vars)))
colnames(or_yn_perc) <- paste(or_yn_vars, "perc", sep = "_")

# Table for yes or no
or_yn_use <- or_yn_perc
colnames(or_yn_use) <- paste(or_yn_vars, "used", sep = "_")

# Compute percent: # days with treatment/# los in hospital
treatment_percent <- function (x) (apply(x, 1, sum, na.rm=T)*100/pi.data$los)

# Get the entire OR data
or_data <- select(pi.data, starts_with("periop"))

for (i in 1:length(or_yn_vars)) {
  # Select a specific treatment type one by one
  or.dt <- dplyr::select(or_data, contains(gsub("periop", "", or_yn_vars[i])))
  
  # Calculate percent of los with each type of treatment
  or_yn_perc[,i] <- treatment_percent(or.dt)
  
  # Check if there is use of a specific treatment type.
  or_yn_use[,i] <- apply(!is.na(or.dt), 1, any)
}

summary(or_yn_perc)

summary(or_yn_use)
```

#### Analyze periop data missing values

```{r}
periop_yn_agg <- cbind(or_yn_use, or_yn_perc)
write.csv(periop_yn_agg, "Analysis/periop_yn_agg.csv")
# get show the missing information matrix           
mdf <- missing_data.frame(periop_agg) 
#show(mdf)
image(mdf)
```

### Get numeric periop data

```{r periop.numeric}
# get the numeric variables
periop_num_vars <- setdiff(periop_vars, or_yn_vars)

# get variables for minimum values
periop_min_vars <- grep("min", periop_num_vars, value = T)

# table to hold absolute min value for each variable
periop_min <- data.frame(matrix(NA, nrow = nrow(ts_data), ncol = length(periop_min_vars)))
colnames(periop_min) <- periop_min_vars

# Get variables with max values
periop_max_vars <- grep("max", periop_num_vars, value = T)

# table to hold absolute min value for each variable
periop_max <- data.frame(matrix(NA, nrow = nrow(ts_data), ncol = length(periop_max_vars)))
colnames(periop_max) <- periop_max_vars

# Compute absolute minimum for the numeric variables
get.min <- function(x) ifelse(all(is.na(x)), NA, min(x, na.rm = T))

for (i in periop_min_vars) {
  min_tbl <- select(or_data, ends_with(gsub("periop", "",i)))
  #print(c(i,dim(min_tbl)))
  periop_min[[i]] <- apply(min_tbl, 1, get.min)
}

# Compute absolute maximum for the numeric variables
get.max <- function(x) ifelse(all(is.na(x)), NA, max(x, na.rm = T))

for (i in periop_max_vars) {
  max_tbl <- select(or_data, ends_with(gsub("periop", "",i)))
  #print(c(i,gsub("periop_", "",i),names(max_tbl)[1:2]))
  periop_max[[i]] <- apply(max_tbl, 1, get.max)
}


##############################################################
# Empty tables to hold the results for mean values
periop_num_mean <- data.frame(matrix(data=NA, nrow = nrow(ts_data), ncol = length(periop_num_vars)))
colnames(periop_num_mean) <- periop_num_vars

# Compute the mean of min and max for all hospital stay period
avg_value <- function (x) (apply(x, 1, mean, na.rm=T))

for (i in periop_num_vars) {
  # Select one type of treatment/feature at a time
  select_dt <- select(or_data, ends_with(gsub("periop","",i)))

  # Compute the mean across length of stay for each individual
  periop_num_mean[[i]] <- avg_value(select_dt)
}
colnames(periop_num_mean) <- paste0(periop_num_vars, "_avg")

###################################################
# Combine all numeric aggregate tables to one table
periop_num_agg <- cbind(periop_min, periop_max, periop_num_mean)

write.csv(periop_num_agg, "Analysis/periop_numeric_agg.csv")
```


## All data in the file

- Static
- dail binary
- Daily numeric
- daily braden
- Periop
- location
- Anesthesia duration
- operation date
- patient in room duration
- surgical schedule
- surgical service
- sugProcDuration

### Anesthesia duration data

```{r fig.width=10}
anesthesia_data <- dplyr::select(pi.data, starts_with("Anesthesia"))

counts <- function(df) apply(df, 1,  function (x) sum(!is.na(x)))

anesth_duration <- counts(anesthesia_data)
histogram(anesth_duration)
# Plot how many observations each column have
count.plot <- function (dt, ttl)
  {obs.counts <- apply(!is.na(dt), 2, function (x) sum(x, na.rm = T)*100/length(x))
    xbar <- barplot(obs.counts, 
                ylim = c(0,100),
                 names.arg = NA,
                 main = ttl,
                 xlab = "Minutes")
text(xbar, 0, labels = 1:26, pos = 1, xpd = T)
text(xbar, obs.counts, labels = round(obs.counts,2), pos = 3, xpd = T)

box( )
}

ttl = "Percent of individual with Anesthesia"

count.plot(anesthesia_data, ttl)
```
#### Number of individual with no anesthesia use or missing data

```{r}
sum(apply(is.na(anesthesia_data), 1, all))
```

### Surgical Procedure Duration data

*Question: What do the numbers represent in this data?*

```{r fig.width=10}
proc.dur_data <- dplyr::select(pi.data, starts_with("SurgProc"))
procedure_time <- counts(proc.dur_data)
histogram(procedure_time)
ttl = "Surgery duration vs percent of individual"

count.plot(proc.dur_data, ttl)
```

### Patient in room Duration data

*Question: What do the numbers represent in this data?*

```{r fig.width=10}
room.dur_data <- dplyr::select(pi.data, starts_with("PatientInRoom"))

inroom_time <- counts(room.dur_data)
# Plot how many observations each column have
ttl = "In room duration vs percent of individual"

count.plot(room.dur_data, ttl)
```

### Surgical Schedule Procedure data

*Question: What does this data represent?*

```{r fig.width=10}
surg.sch_data <- dplyr::select(pi.data, starts_with("SurgicalSchedule"))
surgical_schedule <- apply(surg.sch_data, 1, paste, collapse='*')
```

#### Number of individual with no Surgical Schedule Procedure data

```{r}
sum(apply(is.na(surg.sch_data), 1, all))
```

### Surgical Service data

*Question: What does this data represent?*

```{r fig.width=10}
surg.serv_data <- dplyr::select(pi.data, starts_with("surgicalservice_"))
surgical_serivice <- apply(surg.serv_data, 1, paste, collapse='*')
```

#### Number of individual with no Surgical Service data

```{r}
sum(apply(is.na(surg.serv_data), 1, all))
```

### Operation date data

*Question: What does this data represent?*

```{r fig.width=10}
library(dplyr)
op.date_data <- dplyr::select(pi.data, dplyr::starts_with("operation"))

counts <- function(df) apply(df, 1,  function (x) sum(!is.na(x)))

op_los <- apply(op.date_data, 1,  function (x) sum(!is.na(x)))
summary(op_los)
## Remove patients with no surgical procedure
no.surgury <- apply(apply(op.date_data, 1, is.na), 2,all)
length(no.surgury)
pi <- static_data$pi[!no.surgury]
# Calculate percent pi and non pi
pi.rate <- round(prop.table(table(static_data$pi))*100,2)
pi.rate
table(pi)


## Remove patients with lost less than 2 days
los = as.Date(pi.data$dischargedt)-as.Date(pi.data$admitdt)
los_2 <-(los >=2)
pi <- pi[los_2]
table(pi)
round(prop.table(table(pi))*100,2)
length(pi)

fltr <- apply(matrix(c(!no.surgury,los_2), ncol=2 ),1 ,  all)
table(fltr)
table(static_data$pi[fltr])

table(pi)
op.data <- data.frame(matrix(c(op_los, fltr), ncol=2))
colnames(op.data) <- c("Operation days","Filter")
```

#### Number of individual with no operation date data

```{r}
sum(apply(is.na(op.date_data), 1, all))
```
### Merge all operation related data

```{r}
op_agg_data <- cbind(anesthesia_data, proc.dur_data, room.dur_data, surg.sch_data, surg.serv_data)

write.csv(op_agg_data, "Analysis/surgical_data_agg.csv")
surg_data_agg <- read.csv("Analysis/surgical_data_agg.csv")

surgical_data <- data.frame(cbind(op_los, fltr,anesth_duration,
                       procedure_time,inroom_time,surgical_serivice, surgical_schedule))

surgical_data$encounterid <- pi.data$encounterid


write.csv(surgical_data, "Analysis/surgical_data_processed.csv")

table(static_data$pi[as.logical(surgical_data$fltr)])
table(static_data$pi[fltr])
cor(surgical_data[,3:5])
corrplot::corrplot(cor(surgical_data[,3:5]), addCoef.col = "white")

```

```{r}
library("openxlsx")
asa_data <- read.xlsx("Analysis/IP_op_ASA_data.xlsx")

asa_na <- apply(!apply(asa_data[,3:28], 1, is.na), 2, sum)

asa_data$asa_mean <- apply(asa_data[,3:28], 1, mean, na.rm=TRUE)
asa_data$asa_min <- apply(asa_data[,3:28], 1, min, na.rm=TRUE)
asa_data$asa_min[asa_data$asa_min == Inf] <- -1
asa_data$asa_max <- apply(asa_data[,3:28],1, max, na.rm=TRUE)
asa_data$asa_max[asa_data$asa_max == -Inf] <- -1
asa_data$asa_sd<- apply(asa_data[,3:28],1, sd, na.rm=TRUE)

asa_data_agg <- merge(pi.data[,1:2], asa_data[,c(1,2,29:32)], by=c("encounterid")) #,"encounterid"))
#asa_data_agg <- asa_data_agg %>% dplyr::select(1,2,29,30,31,32)
# SD has too many NA values, so we will drop it.
asa_data_agg <- asa_data_agg[,-7]
write.csv(asa_data_agg, "Analysis/asa_data_agg.csv")

summary(asa_data_agg[,4:6])
```


